---
title: "Full Preregistered analysis"
output: html_document
date: "2026-08-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Prepare dataframe
#### Import dataframe
```{r}
data <- read.csv("Coded_Response_Data.csv")
data$movement <- as.logical(data$movement)
data$colour_change <- as.logical(data$colour_change)
data$valid <- as.logical(data$valid)
```
#### Clean dataframe: 
remove rows where animation misses frames, where report does not feature geometric shapes 
```{r}
frame_cols <- c("frame_1_duration_ms", "frame_2_duration_ms",
                "frame_3_duration_ms", "frame_4_duration_ms")
data <- data[complete.cases(data[, frame_cols]), ]
data <- data[data$valid == TRUE, ]
```

#### Separate illusion, baseline, phi conditions 
```{r}
illusion_condition <- data[data$condition == "illusion_2frames", ]
baseline_condition <- data[data$condition == "baseline_2frames", ]
phi_condition <- data[data$condition == "phi_2frames", ]
illusion_trials <- nrow(illusion_condition)
baseline_trials <- nrow(baseline_condition)
phi_trials <- nrow(phi_condition)
```

## Hypotheses
All hypotheses are tested with proportions z tests, with two-sided alternative hypotheses. 

### Colour phi report task: comparisons between illusion & baseline conditions 
#### H1 [motion reports]: proportions of motion codes.  
```{r}
illusion_motion <- sum(illusion_condition$movement == TRUE)
baseline_motion <- sum(baseline_condition$movement == TRUE)

h1 <- prop.test(x = c(illusion_motion, baseline_motion),
                n = c(illusion_trials, baseline_trials),
                correct = FALSE)
print(h1)
```
```{r}
h1$conf.int
```


#### H2 [colour change reports]: proportions of colour change codes 
```{r}
illusion_change <- illusion_condition[illusion_condition$colour_change == TRUE, ]
baseline_change <- baseline_condition[baseline_condition$colour_change == TRUE, ]
illusion_change_trials = nrow(illusion_change)
baseline_change_trials = nrow(baseline_change)

h2 <- prop.test(x = c(illusion_change_trials, baseline_change_trials),
                n = c(illusion_trials, baseline_trials),
                correct = FALSE)
print(h2)
```

#### H3 [discrete colour change reports]: proportions of dicrete codes, out of colour change responses only 
 [This hypothesis cannot be tested because there are no trials in the baseline condition reporting colour change at all.] 
```{r}
illusion_discrete <- sum(illusion_change$change_type == "Discrete")
baseline_discrete <- sum(baseline_change$change_type == "Discrete")

# h3 <- prop.test(x = c(illusion_discrete, baseline_discrete),
#                n = c(illusion_change_trials, baseline_change_trials),
#                correct = FALSE)
#print(h3)

```

#### H4 [shape number perception]: proportions of single shape reports out of those reporting one or two 
```{r}
illusion_valid <- illusion_condition[illusion_condition$shape_number %in% c("1", "2"), ]
baseline_valid <- baseline_condition[baseline_condition$shape_number %in% c("1", "2"), ]
illusion_valid_trials <- nrow(illusion_valid)
baseline_valid_trials <- nrow(baseline_valid)

illusion_one_shape <- sum(illusion_valid$shape_number == "1")
baseline_one_shape <- sum(baseline_valid$shape_number == "1")

h4 <- prop.test(x = c(illusion_one_shape, baseline_one_shape),
                n = c(illusion_valid_trials, baseline_valid_trials),
                correct = FALSE)
print(h4)
```

#### H5 [frame sequence type]: proportions of choosing four frame sequences
```{r}
illusion_matched_motion <- illusion_condition[illusion_condition$matched_stimuli %in% c("4frames_gradual",
                                                                                        "4frames_discrete"), ]
baseline_matched_motion <- baseline_condition[baseline_condition$matched_stimuli %in% c("4frames_gradual",
                                                                                        "4frames_discrete"), ]
 
illusion_matched_motion_trials <- nrow(illusion_matched_motion)
baseline_matched_motion_trials <- nrow(baseline_matched_motion)

h5 <- prop.test(x = c(illusion_matched_motion_trials, baseline_matched_motion_trials),
                n = c(illusion_trials, baseline_trials),
                correct = FALSE)
print(h5)
```


#### H6 [frame sequence colour change]: proportions of choosing dicrete change frame sequence, among those choosing continous ones
```{r}
illusion_matched_discrete <- sum(illusion_matched_motion$matched_stimuli == "4frames_discrete")
baseline_matched_discrete <- sum(baseline_matched_motion$matched_stimuli == "4frames_discrete")

h6 <- prop.test(x = c(illusion_matched_discrete, baseline_matched_discrete),
                n = c(illusion_matched_motion_trials, baseline_matched_motion_trials),
                correct = FALSE)
print(h6)
```

### Phi motion report task: comparisons between illusion & phi motion conditions.
#### H7 [motion reports, comparison with phi motion]: proportions of motion codes.  
```{r}
phi_motion <- sum(phi_condition$movement == TRUE)

h7 <- prop.test(x = c(illusion_motion, phi_motion),
                n = c(illusion_trials, phi_trials),
                correct = FALSE)
print(h7)
```

#### H8 [shape number perception, comparison with phi motion]: proportions of single shape reports out of those reporting one or two 
```{r}
phi_valid <- phi_condition[phi_condition$shape_number %in% c("1", "2"), ]
phi_valid_trials <- nrow(phi_valid)
phi_one_shape <- sum(phi_valid$shape_number == "1")


h8 <- prop.test(x = c(illusion_one_shape, phi_one_shape),
                n = c(illusion_valid_trials, phi_valid_trials),
                correct = FALSE)
print(h8)
```

#### H9 [frame sequence type, comparison with phi motion]: proportions of choosing four frame sequences
```{r}
phi_matched_motion <- phi_condition[phi_condition$matched_stimuli %in% c("motion_4frames",
                                                                                        "discrete_4frames"), ]
phi_matched_motion_trials <- nrow(phi_matched_motion)


h9 <- prop.test(x = c(illusion_matched_motion_trials, phi_matched_motion_trials),
                n = c(illusion_trials, phi_trials),
                correct = FALSE)
print(h9)
```

## Corrections for multiple comparisons (?)
#### H1 and H7 [motion reports]
```{r}
raw_p <- c(illusion_vs_baseline = h1$p.value,
           illusion_vs_phi = h7$p.value)

adjusted_p <- p.adjust(raw_p, method = "holm")
adjusted_p
```
#### H4 and H8 [shape number perception]
```{r}
raw_p <- c(illusion_vs_baseline = h4$p.value,
           illusion_vs_phi = h8$p.value)

adjusted_p <- p.adjust(raw_p, method = "holm")
adjusted_p
```
#### H5 and H9 [frame sequence type]
```{r}
raw_p <- c(illusion_vs_baseline = h5$p.value,
           illusion_vs_phi = h9$p.value)

adjusted_p <- p.adjust(raw_p, method = "holm")
adjusted_p
```

